Omix

### 2019-04-10 omix/omix.Rmd
updated Rmd w/ mutation DepMap/TCGA comparison

```{r}
library(circlize)
library(ComplexHeatmap)
library(crunch)
library(ggpubr)
library(ggrepel)
library(ggsignif)
library(kableExtra)
library(naniar)
library(openxlsx)
library(reshape2)
library(tidyverse)
library(VennDiagram)

@@

```{r}
unfactorize <- function(df){
  for(i in which(sapply(df, class) == "factor")) {
    df[[i]] = as.character(df[[i]])
  }
  return(df)
}
```

@@

DepMap 19Q1 update fixed almost all of the inconsistencies of the Primary.Disease column, though there were come capitalization inconsistencies.
```{r}
ccl_info <- read.delim("./data/DepMap/DepMap-2019q1-celllines_v2_20190305.csv", sep = ",", header = TRUE, na.strings = c("", NA))

@@

ccl_histologies <- subset(ccl_histologies, !(Broad_ID %in% auc_na_histologies$Broad_ID))
ccl_histologies <- unfactorize(ccl_histologies)
# Bind DepMap histologies and curated histologies
ccl_histologies <- bind_rows(ccl_histologies, auc_na_histologies)
```

@@

Mutation calls (coding region, germline filtered):
```{r, eval=FALSE}
maf_raw <- read.delim("./../data_munging/19Q1/depmap_19Q1_mutation_calls_20190305.csv.gz", header = TRUE, sep = ",", row.names = 1)
# Harmonize ID column name
colnames(maf_raw)[colnames(maf_raw) == "DepMap_ID"] <- "Broad_ID"
# Select columns
maf_df <- subset(maf_raw, select = c("Hugo_Symbol", "Broad_ID", "Variant_Classification", "Reference_Allele", "Tumor_Seq_Allele1", "Genome_Change"))
maf_df$Reference_Allele <- ifelse(maf_df$Reference_Allele == "-", 0, nchar(as.character(maf_df$Reference_Allele)))
maf_df$Tumor_Seq_Allele1 <- nchar(as.character(maf_df$Tumor_Seq_Allele1))
maf_df <- maf_df[maf_df$Variant_Classification != "Silent",]
# Add Mutation_Status column
maf_df$Mutation_Status_Nonsilent <- "Mutant"

@@

DepMap WES CN Data:
```{r, eval=FALSE}
cn <- read.delim("./../data_munging/19Q1/public_19Q1_gene_cn_20190305.csv.gz", sep = ",", check.names = FALSE, header = TRUE)
# Remove Entrez gene IDs from colnames, format is "Hugo_Symbol (Entrez ID)"
colnames(cn) <- gsub(" .*", "", colnames(cn))
colnames(cn)[1] <- "Broad_ID"
# Melt
cn_melt <- melt(data = cn, id.vars = "Broad_ID", measure.vars = colnames(cn[2:ncol(cn)]), variable.name = "Hugo_Symbol", value.name = "CN")
saveRDS(cn_melt, "./../data_munging/19Q1/cn_melt_19Q1.rds", compress = "xz")
```

19Q1, CCLE RNAseq gene expression data (log2(TPM+1)):
```{r, eval=FALSE}
ge_19Q1 <- read.delim("./../../htscreens_giant_files/CCLE_depMap_19Q1_TPM_20190305.csv.gz", header = TRUE, sep = ",", check.names = FALSE)
# Remove Entrez gene IDs from colnames, format is "Hugo_Symbol (Entrez ID)"
colnames(ge_19Q1) <- gsub(" .*", "", colnames(ge_19Q1))
colnames(ge_19Q1)[1] <- "Broad_ID"

# For Denise's GCN
# pgx_ccle_ge <- merge(select(ccl_info, source, source_ccl_name, source_histology, sample_ID, CCLE_ccl_name, Broad_ID), ge_19Q1, by = "Broad_ID", all = TRUE)
# saveRDS(pgx_ccle_ge, "./data_munging/19Q1/pgx_ccle_ge.rds", compress = "xz")

# Melt
ge_melt_19Q1 <- melt(ge_19Q1, id.vars = "Broad_ID", measure.vars = colnames(ge_19Q1)[2:ncol(ge_19Q1)], variable.name = "Hugo_Symbol", value.name = "TPM")
# ge_melt_19Q1 <- unique(ge_melt_19Q1_all)
saveRDS(ge_melt_19Q1, "./../data_munging/19Q1/ge_melt_19Q1.rds", compress = "xz")
```

Load omics RDS:
```{r}
maf_df <- readRDS("./../data_munging/19Q1/maf_df_19Q1.rds")
cn_melt <- readRDS("./../data_munging/19Q1/cn_melt_19Q1.rds")
ge_melt <- readRDS("./../data_munging/19Q1/ge_melt_19Q1.rds")
```

@@

```{r, eval=TRUE}
# Sample info from figshare
crispr_meta <- read.delim("./data/DepMap/sample_info_20190306.csv", sep = ",", header = TRUE, na.strings = c("", NA))
colnames(crispr_meta)[colnames(crispr_meta) == "CCLE_name"] <- "CCLE_Name"

@@

```{r, eval=FALSE}
ccle_poz <- read.delim("./data/pozdeyev2016/CCLE_drug_response_metrics.csv.gz", sep = ",")
# ccle_test <- merge(select(ccle_poz, Drug_name, Cell_line, AUC_IC50), select(filter(auc, Database == "CCLE"), Drug_name, Cell_line, AUC_IC50), by = c("Drug_name", "Cell_line"), suffixes = c("_unmatched", "_matched_all_pairs"), all = TRUE)

@@

```{r}
auc_all <- read.delim("./data/pozdeyev2016/merged_auc_v20190408.5.csv.gz", sep = ",", check.names = FALSE)

@@

```{r, eval=FALSE}
# Data subsetting for easier tracking throughout plots
pgx_ccls <- filter(auc_all, WasDrugScreened == 1)
crispr_ccls <- filter(auc_all, WasCRISPRScreened == 1)
unscreened_ccls <- filter(auc_all, WasCRISPRScreened == 0 & WasDrugScreened == 0)
ge_ccls <- filter(auc_all, HasGeneExpressionData == 1)
cn_ccls <- filter(auc_all, HasCopyNumberData == 1)
mut_ccls <- filter(auc_all, HasMutationData == 1)
ccle_ccls <- filter(auc_all, InCCLE == 1)
ctrp_ccls <- filter(auc_all, InCTRP == 1)
gdsc_ccls <- filter(auc_all, InGDSC == 1)

@@

Data wrangling and processing:
```{r ge_heatmap_data, eval=FALSE}
# DepMap gene expression
ge_depmap <- read.delim("./../../htscreens_giant_files/omix/GE/GE_DepMap_filter.csv.gz", sep = ",", header = TRUE, check.names = FALSE)
colnames(ge_depmap)[colnames(ge_depmap) == "cancer"] <- "TCGA_classification"
# Mean center scale
ge_depmap[, 4:ncol(ge_depmap)] <- apply(ge_depmap[, 4:ncol(ge_depmap)], MARGIN = 2, FUN = scale, center = TRUE, scale = TRUE)
# Melt
ge_depmap_melt <- melt(ge_depmap, id.vars = c("sample_ID", "TCGA_classification"), measure.vars = colnames(ge_depmap)[4:length(colnames(ge_depmap))], variable.name = "Gene", value.name = "TPM_scaled")
# Average scaled TPM per gene per TCGA classification
ge_depmap_avg <- ge_depmap_melt %>% group_by(TCGA_classification, Gene) %>% summarise(Mean_TPM = mean(TPM_scaled))
# Make wide
ge_depmap_wide <- dcast(data = ge_depmap_avg, formula = Gene ~ TCGA_classification, fun.aggregate = mean)
rownames(ge_depmap_wide) <- ge_depmap_wide$Gene
ge_depmap_wide$Gene <- NULL
colnames(ge_depmap_wide) <- paste(colnames(ge_depmap_wide), "DepMap", sep = "_")

@@

Pseudo-figure caption: This is comparison of gene expression profiles between DepMap cancer cell lines and TCGA tumor samples, both classified by TCGA project. In each dataset and for each TCGA classification, gene expression (in TPM) for each of the 5,000 most variable genes shared between the two datasets were averaged across samples. The dataset averages per TCGA classification were correlated against each other, and these Pearson correlations (r) between TCGA and DepMap gene expression feature sets are shown in the heatmap.

@@

Data wrangling and processing:

**DO NOT RUN UNTIL ADDED TCGA TO CN AND GE DATASETS**
```{r cn_heatmap_data, eval=FALSE}
# TCGA copy number
cn_tcga <- read.delim("./data/CN/CN_GDC_filter.csv.gz", sep = ",", header = TRUE, check.names = FALSE)
colnames(cn_tcga)[colnames(cn_tcga) == "cancer"] <- "TCGA_classification"
# Melt
cn_tcga_melt <- melt(cn_tcga, id.vars = c("sample_ID", "TCGA_classification"), measure.vars = colnames(cn_tcga)[4:length(colnames(cn_tcga))], variable.name = "Gene", value.name = "CN")
# Average scaled TPM per gene per TCGA classification
cn_tcga_avg <- cn_tcga_melt %>% group_by(TCGA_classification, Gene) %>% summarise(Mean_CN = mean(CN))
# Make wide
cn_tcga_wide <- dcast(data = cn_tcga_avg, formula = Gene ~ TCGA_classification, fun.aggregate = mean)
rownames(cn_tcga_wide) <- cn_tcga_wide$Gene
cn_tcga_wide$Gene <- NULL
colnames(cn_tcga_wide) <- paste(colnames(cn_tcga_wide), "TCGA", sep = "_")

## 18Q4

@@

# Correlate all against all (18Q4)
cn_wide <- bind_cols(cn_depmap_wide, cn_tcga_wide)
cn_corr <- cn_wide %>% as.matrix %>% cor(method = "spearman") %>% as.data.frame %>% rownames_to_column(var = "var1") %>% gather(var2, value, -var1)
# Filter self-comparisons and eliminate duplicate correlations
cn_corr <- filter(cn_corr, !(grepl("DepMap", var1) & grepl("DepMap", var2)) & !(grepl("TCGA", var1) & grepl("TCGA", var2)) & grepl("TCGA", var2))
# Data management
colnames(cn_corr) <- c("DepMap", "TCGA", "r")
cn_corr <- data.frame(apply(cn_corr, 2, function(x) gsub("_TCGA|_DepMap","", x)))
cn_corr$r <- as.numeric(as.character(cn_corr$r))
# Set up coprrelation dataframe for plotting
cn_corr_wide <- dcast(cn_corr, DepMap ~ TCGA)
rownames(cn_corr_wide) <- cn_corr_wide$DepMap
cn_corr_wide$DepMap <- NULL

## 19Q1

# DepMap copy number (19Q1)
# Filter for genes and TCGA classifications in TCGA CN dataset
cn_filt <- filter(cn_melt, Hugo_Symbol %in% unique(cn_tcga_melt$Gene) & TCGA_classification %in% unique(cn_tcga_melt$TCGA_classification))
colnames(cn_filt)[colnames(cn_filt) == "Hugo_Symbol"] <- "Gene"
# Assign CN levels
cn_19Q1$CN <- with(cn_19Q1, ifelse(CN > 0.3, 1, ifelse(CN < -0.3, -1, 0)))
# Average CN levels per gene per TCGA classification
cn_19Q1_avg <- cn_19Q1 %>% group_by(TCGA_classification, Gene) %>% summarise(Mean_CN = mean(CN))
# Make wide
cn_19Q1_wide <- dcast(data = cn_19Q1_avg, formula = Gene ~ TCGA_classification, fun.aggregate = mean)
rownames(cn_19Q1_wide) <- cn_19Q1_wide$Gene
cn_19Q1_wide$Gene <- NULL
colnames(cn_19Q1_wide) <- paste(colnames(cn_19Q1_wide), "DepMap", sep = "_")

# Correlate all against all (19Q1)
cn_wide_19Q1 <- bind_cols(cn_19Q1_wide, cn_tcga_wide)
cn_corr_19Q1 <- cn_wide_19Q1 %>% as.matrix %>% cor(method = "spearman") %>% as.data.frame %>% rownames_to_column(var = "var1") %>% gather(var2, value, -var1)
# Filter self-comparisons and eliminate duplicate correlations
cn_corr_19Q1 <- filter(cn_corr_19Q1, !(grepl("DepMap", var1) & grepl("DepMap", var2)) & !(grepl("TCGA", var1) & grepl("TCGA", var2)) & grepl("TCGA", var2))
# Data management
colnames(cn_corr_19Q1) <- c("DepMap", "TCGA", "r")
cn_corr_19Q1 <- data.frame(apply(cn_corr_19Q1, 2, function(x) gsub("_TCGA|_DepMap","", x)))
cn_corr_19Q1$r <- as.numeric(as.character(cn_corr_19Q1$r))
# Set up coprrelation dataframe for plotting
cn_corr_19Q1_wide <- dcast(cn_corr_19Q1, DepMap ~ TCGA)
rownames(cn_corr_19Q1_wide) <- cn_corr_19Q1_wide$DepMap
cn_corr_19Q1_wide$DepMap <- NULL
```

Pseudo-figure caption: This is comparison of thresholded copy number profiles between DepMap cancer cell lines and TCGA tumor samples, both classified by TCGA project. Levels are -1 for copuy number loss, 0 for neutral, and 1 for gain. In each dataset and for each TCGA classification, these numeric copy number levels across all genes shared between the two datasets were averaged, and the datasets were correlated against each other. Pearson correlations (r) between TCGA and DepMap copy number feature sets are shown in the heatmap.

Make heatmaps:
```{r cn_heatmaps, eval=FALSE, fig.height=4, fig.width=3.75}
col_fun <- colorRamp2(c(-0.5, 0, 0.5, 1), c("black", "red3", "yellow", "white"))
cn_ht_legend_list <- list(legend_direction = "horizontal", title_position = "lefttop", legend_width = unit(1.5, "in"), border = "black")

# 18Q4
cn_ht4order <- Heatmap(cn_corr_wide, cluster_rows = FALSE)
cn_ht <- Heatmap(cn_corr_wide, 
                 name = "Expression correlation",
                 col = col_fun, 
                 rect_gp = gpar(col = "black"),
                 heatmap_legend_param = cn_ht_legend_list,
                 row_order = rev(column_order(cn_ht4order)), 
                 show_column_dend = FALSE, 
                 cluster_rows = FALSE, 
                 row_names_side = "left",
                 row_title = "DepMap cancer cell lines",
                 column_title = "TCGA tumor samples", 
                 column_title_side = "bottom")
# png(filename = "./figures/cn_depmap_tcga_heatmap_spearman.png", width = 7.5, height = 8, units = "in", res = 500)
draw(cn_ht, heatmap_legend_side = "top")
# dev.off()

# 19Q1
cn_19Q1_ht4order <- Heatmap(cn_corr_19Q1_wide, cluster_rows = FALSE)

cn_19Q1_ht <- Heatmap(cn_corr_19Q1_wide, 
                 name = "Expression correlation",
                 col = col_fun, 
                 rect_gp = gpar(col = "black"),
                 heatmap_legend_param = cn_ht_legend_list,
                 row_order = rev(column_order(cn_19Q1_ht4order)), 
                 show_column_dend = FALSE, 
                 cluster_rows = FALSE, 
                 row_names_side = "left",
                 row_title = "DepMap cancer cell lines",
                 column_title = "TCGA tumor samples", 
                 column_title_side = "bottom")
# png(filename = "./figures/cn_depmap_19Q1_tcga_heatmap_spearman.png", width = 7.5, height = 8, units = "in", res = 500)
draw(cn_19Q1_ht, heatmap_legend_side = "top")
# dev.off()
```

@@

```{r mut_heatmap_data, eval=FALSE}

```

>>>>

```{r load_libraries}
library(circlize)
library(ComplexHeatmap)
library(crunch)
library(ggpubr)
library(ggrepel)
library(ggsignif)
library(kableExtra)
library(naniar)
library(openxlsx)
library(reshape2)
library(tidyverse)
library(VennDiagram)

@@

```{r functions}
unfactorize <- function(df){
  for(i in which(sapply(df, class) == "factor")) {
    df[[i]] = as.character(df[[i]])
  }
  return(df)
}
```

@@

DepMap 19Q1 update fixed almost all of the inconsistencies of the Primary.Disease column, though there were come capitalization inconsistencies.
```{r ccl_info_processing, eval=FALSE}
ccl_info <- read.delim("./data/DepMap/DepMap-2019q1-celllines_v2_20190305.csv", sep = ",", header = TRUE, na.strings = c("", NA))

@@

ccl_histologies <- subset(ccl_histologies, !(Broad_ID %in% auc_na_histologies$Broad_ID))
ccl_histologies <- unfactorize(ccl_histologies)
# Bind DepMap histologies and curated histologies
ccl_histologies <- bind_rows(ccl_histologies, auc_na_histologies)
write.csv(ccl_histologies, "./data/ccl_histologies_metadata.csv")
```

Load cell line metadata file:
```{r}
ccl_histologies <- read.delim("./data/ccl_histologies_metadata.csv", sep = ",", header = TRUE, row.names = 1)
```

@@

Mutation calls (coding region, germline filtered):
```{r mutation, eval=FALSE}
maf_raw <- read.delim("./../data_munging/19Q1/depmap_19Q1_mutation_calls_20190305.csv.gz", header = TRUE, sep = ",", row.names = 1)
# Harmonize ID column name
colnames(maf_raw)[colnames(maf_raw) == "DepMap_ID"] <- "Broad_ID"
colnames(maf_raw)[colnames(maf_raw) == "Hugo_Symbol"] <- "Gene"
# Merge CCL metadata
maf_raw <- merge(select(ccl_histologies, Broad_ID, TCGA_classification), maf_raw, by = "Broad_ID", all.y = TRUE)
# Select columns
maf_df <- subset(maf_raw, select = c("Gene", "Broad_ID", "TCGA_classification", "Variant_Classification", "Reference_Allele", "Tumor_Seq_Allele1", "Genome_Change"))
maf_df$Reference_Allele <- ifelse(maf_df$Reference_Allele == "-", 0, nchar(as.character(maf_df$Reference_Allele)))
maf_df$Tumor_Seq_Allele1 <- nchar(as.character(maf_df$Tumor_Seq_Allele1))
maf_df <- maf_df[maf_df$Variant_Classification != "Silent",]
# Add Mutation_Status column
maf_df$Mutation_Status <- 1

@@

DepMap WES CN Data:
```{r copy_number, eval=FALSE}
cn <- read.delim("./../data_munging/19Q1/public_19Q1_gene_cn_20190305.csv.gz", sep = ",", check.names = FALSE, header = TRUE)
# Remove Entrez gene IDs from colnames, format is "Hugo_Symbol (Entrez ID)"
colnames(cn) <- gsub(" .*", "", colnames(cn))
colnames(cn)[1] <- "Broad_ID"
# Merge CCL metadata with TCGA labels
cn <- merge(select(ccl_histologies, Broad_ID, TCGA_classification), cn, by = "Broad_ID", all.y = TRUE)
# Melt
cn_melt <- melt(data = cn, id.vars = c("Broad_ID", "TCGA_classification"), measure.vars = colnames(cn[3:ncol(cn)]), variable.name = "Gene", value.name = "CN")
saveRDS(cn_melt, "./../data_munging/19Q1/cn_melt_19Q1.rds", compress = "xz")
```

19Q1, CCLE RNAseq gene expression data (log2(TPM+1)):
```{r gene_expression, eval=FALSE}
ge <- read.delim("./../../htscreens_giant_files/CCLE_depMap_19Q1_TPM_20190305.csv.gz", header = TRUE, sep = ",", check.names = FALSE)
# Remove Entrez gene IDs from colnames, format is "Hugo_Symbol (Entrez ID)"
colnames(ge) <- gsub(" .*", "", colnames(ge))
colnames(ge)[1] <- "Broad_ID"
# Merge CCL metadata with TCGA labels
ge <- merge(select(ccl_histologies, Broad_ID, TCGA_classification), ge, by = "Broad_ID", all.y = TRUE)
# Melt
ge_melt <- melt(ge, id.vars = c("Broad_ID", "TCGA_classification"), measure.vars = colnames(ge)[3:ncol(ge)], variable.name = "Gene", value.name = "TPM")
saveRDS(ge_melt, "./../data_munging/19Q1/ge_melt_19Q1.rds", compress = "xz")
```

Load omics RDS:
```{r load_omics_rds}
maf_df <- readRDS("./../data_munging/19Q1/maf_df_19Q1.rds")
cn_melt <- readRDS("./../data_munging/19Q1/cn_melt_19Q1.rds")
ge_melt <- readRDS("./../data_munging/19Q1/ge_melt_19Q1.rds")
```

@@

```{r crispr, eval=TRUE}
# Sample info from figshare
crispr_meta <- read.delim("./data/DepMap/sample_info_20190306.csv", sep = ",", header = TRUE, na.strings = c("", NA))
colnames(crispr_meta)[colnames(crispr_meta) == "CCLE_name"] <- "CCLE_Name"

@@

```{r pozdeyev_processing, eval=FALSE}
ccle_poz <- read.delim("./data/pozdeyev2016/CCLE_drug_response_metrics.csv.gz", sep = ",")
# ccle_test <- merge(select(ccle_poz, Drug_name, Cell_line, AUC_IC50), select(filter(auc, Database == "CCLE"), Drug_name, Cell_line, AUC_IC50), by = c("Drug_name", "Cell_line"), suffixes = c("_unmatched", "_matched_all_pairs"), all = TRUE)

@@

```{r pozdeyev_summaries}
auc_all <- read.delim("./data/pozdeyev2016/merged_auc_v20190408.5.csv.gz", sep = ",", check.names = FALSE)

@@

```{r exploratory_plots, eval=FALSE}
# Data subsetting for easier tracking throughout plots
pgx_ccls <- filter(auc_all, WasDrugScreened == 1)
crispr_ccls <- filter(auc_all, WasCRISPRScreened == 1)
unscreened_ccls <- filter(auc_all, WasCRISPRScreened == 0 & WasDrugScreened == 0)
ge_ccls <- filter(auc_all, HasGeneExpressionData == 1)
cn_ccls <- filter(auc_all, HasCopyNumberData == 1)
mut_ccls <- filter(auc_all, HasMutationData == 1)
ccle_ccls <- filter(auc_all, InCCLE == 1)
ctrp_ccls <- filter(auc_all, InCTRP == 1)
gdsc_ccls <- filter(auc_all, InGDSC == 1)

@@

Data wrangling and processing:
```{r ge_heatmap_data, eval=FALSE, warning=FALSE}
# DepMap gene expression
ge_depmap <- read.delim("./../../htscreens_giant_files/omix/GE/GE_DepMap_filter.csv.gz", sep = ",", header = TRUE, check.names = FALSE)
colnames(ge_depmap)[colnames(ge_depmap) == "cancer"] <- "TCGA_classification"
# Mean center scale
ge_depmap[, 4:ncol(ge_depmap)] <- apply(ge_depmap[, 4:ncol(ge_depmap)], MARGIN = 2, FUN = scale, center = TRUE, scale = TRUE)
# Melt
ge_depmap_melt <- melt(ge_depmap, id.vars = c("sample_ID", "TCGA_classification"), measure.vars = colnames(ge_depmap)[4:length(colnames(ge_depmap))], variable.name = "Gene", value.name = "TPM_scaled")
# Average scaled TPM per gene per TCGA classification
ge_depmap_avg <- ge_depmap_melt %>% group_by(TCGA_classification, Gene) %>% summarise(Mean_TPM = mean(TPM_scaled))
# Make wide
ge_depmap_wide <- dcast(data = ge_depmap_avg, formula = Gene ~ TCGA_classification, fun.aggregate = mean)
rownames(ge_depmap_wide) <- ge_depmap_wide$Gene
ge_depmap_wide$Gene <- NULL
colnames(ge_depmap_wide) <- paste(colnames(ge_depmap_wide), "DepMap", sep = "_")

@@

Pseudo-figure caption: This is comparison of gene expression profiles between DepMap cancer cell lines and TCGA tumor samples, both classified by TCGA project. In each dataset and for each TCGA classification, gene expression (in TPM) for each of the 5,000 most variable genes shared between the two datasets were averaged across samples. The dataset averages per TCGA classification were correlated against each other, and these Pearson/Spearman correlations (r) between TCGA and DepMap gene expression feature sets are shown in the heatmap.

@@

Data wrangling and processing:
```{r cn_heatmap_data, eval=FALSE, warning=FALSE}
# TCGA copy number
cn_tcga <- read.delim("./data/CN/CN_GDC_filter.csv.gz", sep = ",", header = TRUE, check.names = FALSE)
colnames(cn_tcga)[colnames(cn_tcga) == "cancer"] <- "TCGA_classification"
# Melt
cn_tcga_melt <- melt(cn_tcga, id.vars = c("sample_ID", "TCGA_classification"), measure.vars = colnames(cn_tcga)[4:length(colnames(cn_tcga))], variable.name = "Gene", value.name = "CN")
# Average scaled TPM per gene per TCGA classification
cn_tcga_avg <- cn_tcga_melt %>% group_by(TCGA_classification, Gene) %>% summarise(Mean_CN = mean(CN))
# Make wide
cn_tcga_wide <- dcast(data = cn_tcga_avg, formula = Gene ~ TCGA_classification, fun.aggregate = mean)
rownames(cn_tcga_wide) <- cn_tcga_wide$Gene
cn_tcga_wide$Gene <- NULL
colnames(cn_tcga_wide) <- paste(colnames(cn_tcga_wide), "TCGA", sep = "_")

@@

# Correlate all against all (18Q4)
cn_wide <- bind_cols(cn_depmap_wide, cn_tcga_wide)
cn_corr <- cn_wide %>% as.matrix %>% cor(method = "pearson") %>% as.data.frame %>% rownames_to_column(var = "var1") %>% gather(var2, value, -var1)
# Filter self-comparisons and eliminate duplicate correlations
cn_corr <- filter(cn_corr, !(grepl("DepMap", var1) & grepl("DepMap", var2)) & !(grepl("TCGA", var1) & grepl("TCGA", var2)) & grepl("TCGA", var2))
# Data management
colnames(cn_corr) <- c("DepMap", "TCGA", "r")
cn_corr <- data.frame(apply(cn_corr, 2, function(x) gsub("_TCGA|_DepMap","", x)))
cn_corr$r <- as.numeric(as.character(cn_corr$r))
# Set up coprrelation dataframe for plotting
cn_corr_wide <- dcast(cn_corr, DepMap ~ TCGA)
rownames(cn_corr_wide) <- cn_corr_wide$DepMap
cn_corr_wide$DepMap <- NULL
```

Pseudo-figure caption: This is comparison of thresholded copy number profiles between DepMap cancer cell lines and TCGA tumor samples, both classified by TCGA project. Levels are -1 for copy number loss, 0 for neutral, and 1 for gain. In each dataset and for each TCGA classification, these numeric copy number levels across all genes shared between the two datasets were averaged, and the datasets were correlated against each other. Pearson/Spearman correlations (r) between TCGA and DepMap copy number feature sets are shown in the heatmap.

Make heatmaps:
```{r cn_heatmaps, eval=FALSE, fig.height=4, fig.width=3.75}
col_fun <- colorRamp2(c(0, 0.25, 0.5, 0.75), c("black", "red3", "yellow", "white"))
cn_ht_legend_list <- list(legend_direction = "horizontal", title_position = "lefttop", legend_width = unit(1.5, "in"), border = "black")

# 18Q4
cn_ht4order <- Heatmap(cn_corr_wide, cluster_rows = FALSE)
cn_ht <- Heatmap(cn_corr_wide, 
                 name = "Copy number correlation",
                 col = col_fun, 
                 rect_gp = gpar(col = "black"),
                 heatmap_legend_param = cn_ht_legend_list,
                 row_order = rev(column_order(cn_ht4order)), 
                 show_column_dend = FALSE, 
                 cluster_rows = FALSE, 
                 row_names_side = "left",
                 row_title = "DepMap cancer cell lines",
                 column_title = "TCGA tumor samples", 
                 column_title_side = "bottom")
# png(filename = "./figures/cn_depmap_tcga_heatmap_pearson.png", width = 7.5, height = 8, units = "in", res = 500)
draw(cn_ht, heatmap_legend_side = "top")
# dev.off()
```

@@

Data wrangling and processing:
```{r mut_heatmap_data, eval=FALSE}
# TCGA
mut_tcga <- read.delim("./data/mutation/mutation_Xena_filter.csv.gz", sep = ",", header = TRUE, check.names = FALSE)
mut_tcga <- unfactorize(mut_tcga)
mut_tcga[, 4:ncol(mut_tcga)] <- apply(mut_tcga[, 4:ncol(mut_tcga)], 2, function(x) ifelse(x > 1, 1, x))
colnames(mut_tcga)[colnames(mut_tcga) == "cancer"] <- "TCGA_classification"
# Melt
mut_tcga_melt <- melt(mut_tcga, id.vars = c("sample_ID", "TCGA_classification"), measure.vars = colnames(mut_tcga)[4:length(colnames(mut_tcga))], variable.name = "Gene", value.name = "Mutation_Status")
# Average CN levels per gene per TCGA classification
mut_tcga_prop <- mut_tcga_melt %>% group_by(TCGA_classification, Gene) %>% summarise(Proportion_Mut = sum(Mutation_Status)/n())
# Make wide
mut_tcga_wide <- dcast(data = mut_tcga_prop, formula = Gene ~ TCGA_classification, fun.aggregate = mean)
rownames(mut_tcga_wide) <- mut_tcga_wide$Gene
mut_tcga_wide$Gene <- NULL
colnames(mut_tcga_wide) <- paste(colnames(mut_tcga_wide), "TCGA", sep = "_")

# DepMap
mut_depmap <- read.delim("./data/mutation/mutation_DepMap_filter.csv.gz", sep = ",", header = TRUE, check.names = FALSE)
mut_depmap <- unfactorize(mut_depmap)
mut_depmap[, 4:ncol(mut_depmap)] <- apply(mut_depmap[, 4:ncol(mut_depmap)], 2, function(x) ifelse(x > 1, 1, x))
colnames(mut_depmap)[colnames(mut_depmap) == "cancer"] <- "TCGA_classification"
# Melt
mut_depmap_melt <- melt(mut_depmap, id.vars = c("sample_ID", "TCGA_classification"), measure.vars = colnames(mut_depmap)[4:length(colnames(mut_depmap))], variable.name = "Gene", value.name = "Mutation_Status")
# Average CN levels per gene per TCGA classification
mut_depmap_prop <- mut_depmap_melt %>% group_by(TCGA_classification, Gene) %>% summarise(Proportion_Mut = sum(Mutation_Status)/n())
# Make wide
mut_depmap_wide <- dcast(data = mut_depmap_prop, formula = Gene ~ TCGA_classification, fun.aggregate = mean)
rownames(mut_depmap_wide) <- mut_depmap_wide$Gene
mut_depmap_wide$Gene <- NULL
colnames(mut_depmap_wide) <- paste(colnames(mut_depmap_wide), "DepMap", sep = "_")

# Correlate all against all (18Q4)
mut_wide <- bind_cols(mut_depmap_wide, mut_tcga_wide)
mut_corr <- mut_wide %>% as.matrix %>% cor(method = "spearman") %>% as.data.frame %>% rownames_to_column(var = "var1") %>% gather(var2, value, -var1)
# Filter self-comparisons and eliminate duplicate correlations
mut_corr <- filter(mut_corr, !(grepl("DepMap", var1) & grepl("DepMap", var2)) & !(grepl("TCGA", var1) & grepl("TCGA", var2)) & grepl("TCGA", var2))
# Data management
colnames(mut_corr) <- c("DepMap", "TCGA", "r")
mut_corr <- data.frame(apply(mut_corr, 2, function(x) gsub("_TCGA|_DepMap","", x)))
mut_corr$r <- as.numeric(as.character(mut_corr$r))
# Set up coprrelation dataframe for plotting
mut_corr_wide <- dcast(mut_corr, DepMap ~ TCGA)
rownames(mut_corr_wide) <- mut_corr_wide$DepMap
mut_corr_wide$DepMap <- NULL
```

Pseudo-figure caption: This is comparison of binary gene-level mutation frequency profiles between DepMap cancer cell lines and TCGA tumor samples, both classified by TCGA project. In each dataset and for each TCGA classification, the proportion of samples containing at least one mutation was calculated for all genes shared between the two datasets, and the proportions were correlated against each other. Pearson/Spearman correlations (r) between TCGA and DepMap copy number feature sets are shown in the heatmap.

Genes in the TCGA mutation dataset were analyzed if they were mutated in >=4% annotated samples and had a mutaiton in at least one cancer cell line from DepMap.

Make heatmaps:
```{r cn_heatmaps, eval=FALSE, fig.height=4, fig.width=3.75}
col_fun <- colorRamp2(c(0, 0.15, 0.35, 0.5), c("black", "red3", "yellow", "white"))
mut_ht_legend_list <- list(legend_direction = "horizontal", title_position = "lefttop", legend_width = unit(1.5, "in"), border = "black")

# 18Q4
mut_ht4order <- Heatmap(mut_corr_wide, cluster_rows = FALSE)
mut_ht <- Heatmap(mut_corr_wide, 
                 name = "Mutation correlation",
                 col = col_fun, 
                 rect_gp = gpar(col = "black"),
                 heatmap_legend_param = mut_ht_legend_list,
                 row_order = rev(column_order(mut_ht4order)), 
                 show_column_dend = FALSE, 
                 cluster_rows = FALSE, 
                 row_names_side = "left",
                 row_title = "DepMap cancer cell lines",
                 column_title = "TCGA tumor samples", 
                 column_title_side = "bottom")
# png(filename = "./figures/mut_depmap_tcga_heatmap_spearman.png", width = 7.5, height = 8, units = "in", res = 500)
draw(mut_ht, heatmap_legend_side = "top")
# dev.off()
```
